\section{Dimensionality}
    
    \begin{definition*}[Dimensionality]
        Given a formula $P$, the coalescence proof search produces a proof in an $n$-dimensional petri net.
        Equivalently, an additively stratified sequent proof $\vdash P$ has $n$ terms at the bottom before any contractions are applied.
        The dimensionality of $P$ is then defined $\dim(P) \defeq n$.
    \end{definition*}

    \begin{example*}
        Borrowing the proof from Example~\ref{example:add-strat-proof}, it has dimensionality 2 since the tree concludes with $\vdash P, P$ before any contractions.
        \begin{prooftree}
            \AxiomC{}
            \RightLabel{$ax$}\UnaryInfC{$\vdash a, \neg a$}
            \RightLabel{$\vee$}\UnaryInfC{$\vdash a \vee \neg a, \neg a$}
            \RightLabel{$\vee$}\UnaryInfC{$\vdash a \vee \neg a, a \vee \neg a$}
            \AxiomC{}
            \RightLabel{$\top$}\UnaryInfC{$\vdash \top$}
            \RightLabel{$w$}\UnaryInfC{$\vdash \top, a \vee \neg a$}
            \RightLabel{$\wedge$}\BinaryInfC{$\vdash (a \vee \neg a) \wedge \top, a \vee \neg a$}
            \AxiomC{}
            \RightLabel{$\top$}\UnaryInfC{$\vdash \top$}
            \RightLabel{$w$}\UnaryInfC{$\vdash \top, (a \vee \neg a) \wedge \top$}
            \RightLabel{$\wedge$}\BinaryInfC{$\vdash (a \vee \neg a) \wedge \top, (a \vee \neg a) \wedge \top$}
            \RightLabel{$c$}\UnaryInfC{$\vdash (a \vee \neg a) \wedge \top$}
        \end{prooftree}
    \end{example*}


    \begin{definition}[Classes of Formulae]\label{sec-ctr:formulae-classes}
        Let $D^i$ be the subclasses of formulae defined as:
        \begin{align*}
            D^1 \quad &\defeq \quad \top \,|\, D^1 \wedge D^1 \,|\, D^* \vee D^1 \\
            D^2 \quad &\defeq \quad D^{i \geq 2} \vee D^2 \,|\, P \vee \neg P \,|\, \ldots
        \end{align*}
        where $D^* \defeq \bigcup_{i \in \mathbb{N}} D_i$, such that $D^n$ is the class of all formulae provable in n dimensions.

        \textit{This definition will be revised later.}
    \end{definition}


    \begin{remark*}[Satisfiability vs Provability]
        For any formula $P$, there exist four distinct classes: \textit{true, false, satisfiably true, satisfiably false}, where satisfiable differs by finding a particular assignment of values to each variable.
        Coalescence searches for a proof of $P$ in \textit{true}, whereas the SAT problem addresses a proof of $P$ in \textit{true}.
    \end{remark*}
    
    
    \begin{proposition*}[Bounds on Dimensionality]
        Given a formula $P$, the following provides a bound its dimensionality:
        \begin{equation*}
            P \in D^n \implies n+1 \leq vars(P)
        \end{equation*}
        Justification is left unaddressed here as the current construction does not emit a `nice' proof.
    \end{proposition*}

    \begin{remark*}
        These methods do not lend themselves to `nice' properties.
        It is possible that in the general case, \textit{finding an exact value for the dimensionality of a given formula $P$ may be equivalent to finding a proof of $P$}, but it remains to be proven.
        Instead, a natural optimisation is applied to the algorithm as follows, giving rise to a different set of classes of dimensionality and more pleasing properties.
    \end{remark*} 

    \begin{definition*}[$\top$-substitution]
         Given a sequent expressing a proof of $Q$, i.e. $\vdash Q, Q, \ldots Q$, the formula $P$ may be equivalently expressed $P[Q \seteq \top]$.
    \end{definition*}

    \begin{example*}
        Consider the formula $P \defeq (a \vee \neg a) \wedge (b \vee \neg b)$.
        The sequent proof $\vdash P$ is large and has dimensionality 3 as follows:
        \begin{center}
            \begin{tabular}{@{}l@{}}
                \begin{varwidth}{\linewidth}
                    \begin{scprooftree}{0.47}
                        \AxiomC{$ \vdash a , a , \neg a $}
                        \UnaryInfC{$ \vdash a , ( a \vee \neg a ) , \neg a $}
                        \AxiomC{$ \vdash a , \neg a , b $}
                        \UnaryInfC{$ \vdash a , \neg a , ( b \vee \neg b ) $}
                        \BinaryInfC{$ \vdash a , \neg a , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) $}
                        \UnaryInfC{$ \vdash a , ( a \vee \neg a ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) $}
                        \AxiomC{$ \vdash a , \neg a , b $}
                        \UnaryInfC{$ \vdash a , \neg a , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash a , ( a \vee \neg a ) , ( b \vee \neg b ) $}
                        \AxiomC{$ \vdash a , b , \neg b $}
                        \UnaryInfC{$ \vdash a , b , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash a , ( b \vee \neg b ) , ( b \vee \neg b ) $}
                        \BinaryInfC{$ \vdash a , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( b \vee \neg b ) $}
                        \BinaryInfC{$ \vdash a , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) $}
                        \UnaryInfC{$ \vdash ( a \vee \neg a ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) $}
                        \AxiomC{$ \vdash a , \neg a , b $}
                        \UnaryInfC{$ \vdash a , \neg a , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash a , ( a \vee \neg a ) , ( b \vee \neg b ) $}
                        \AxiomC{$ \vdash a , b , \neg b $}
                        \UnaryInfC{$ \vdash a , b , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash a , ( b \vee \neg b ) , ( b \vee \neg b ) $}
                        \BinaryInfC{$ \vdash a , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash ( a \vee \neg a ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( b \vee \neg b ) $}
                        \AxiomC{$ \vdash ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , b , \neg b $}
                        \UnaryInfC{$ \vdash ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , b , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( b \vee \neg b ) , ( b \vee \neg b ) $}
                        \BinaryInfC{$ \vdash ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( b \vee \neg b ) $}
                        \BinaryInfC{$ \vdash ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) , ( ( a \vee \neg a ) \wedge ( b \vee \neg b ) ) $}
                    \end{scprooftree}
                \end{varwidth}
            \end{tabular}
        \end{center}
        
        Using substitutions of $\top$ for provable subformulae, the proof $\vdash P$ becomes much more manageable:

        \begin{center}
            \begin{tabular}{@{}l@{}}
                $ A := ( a \vee \neg a ) $
                \begin{varwidth}{\linewidth}
                    \begin{prooftree}
                        \AxiomC{$ \vdash a , \neg a $}
                        \UnaryInfC{$ \vdash a , ( a \vee \neg a ) $}
                        \UnaryInfC{$ \vdash ( a \vee \neg a ) , ( a \vee \neg a ) $}
                    \end{prooftree}
                \end{varwidth}\\ \\

                $ B := ( b \vee \neg b ) $
                \begin{varwidth}{\linewidth}
                    \begin{prooftree}
                        \AxiomC{$ \vdash b , \neg b $}
                        \UnaryInfC{$ \vdash b , ( b \vee \neg b ) $}
                        \UnaryInfC{$ \vdash ( b \vee \neg b ) , ( b \vee \neg b ) $}
                    \end{prooftree}
                \end{varwidth}\\ \\

                \begin{varwidth}{\linewidth}
                    \begin{prooftree}
                        \AxiomC{$ \vdash A , A $}
                        \AxiomC{$ \vdash A , B $}
                        \BinaryInfC{$ \vdash A , ( A \wedge B ) $}
                        \AxiomC{$ \vdash A , B $}
                        \AxiomC{$ \vdash B , B $}
                        \BinaryInfC{$ \vdash B , ( A \wedge B ) $}
                        \BinaryInfC{$ \vdash ( A \wedge B ) , ( A \wedge B ) $}
                    \end{prooftree}
                \end{varwidth}\\
            \end{tabular}
        \end{center}

        While the final step need only be performed with dimensionality 1, the prior steps require dimensionality 2, giving the full formula a dimensionality of 2.
    \end{example*}


    \begin{remark*}
        It is now important to reexamine the coalescence algorithm.

        Storing a token as a \texttt{Tuple} gives an inefficient algorithm due to the implicit ordering.
        By commutativity, given a token $(a_1 \ldots a_n)$, there exist up to $n!$ permutations of $a_i$'s and thus $n!$ equivalent tokens.
        
        Instead, order is abstracted over and a token is stored as a \texttt{Multiset}.
        This gives the storage optimisation exploiting the symmetry of petri nets that expresses commutativity of terms in a sequent.

        This may be taken a step further.
        Consider a token stored as a \texttt{Set} which expresses idempotency of terms in a sequent $\vdash P, P \equiv \, \vdash P$.
    \end{remark*}

    \begin{examples}
        Consider again the formula $(a \vee \neg a) \wedge (b \vee \neg b)$, with a proofs as follow, but instead tracing the values of tokens.
        
        Where tokens are member of the category \texttt{Tuple}, so as to demonstrate the exponential blow-up, both the `minimal' (black) and `saturated' (grey) proofs are as follows:
        \input{proofnet-example-tuple} where $\rightsquigarrow$ is an application of weakening.
        This is notable as this will be the point coalescence will reach for $n \seteq 2$ and a suitably intelligent method for extending dimensions will maintain the proof search so far.
        Both `minimal' and `saturated' proofs are included as each represent the deterministic versus non-deterministic searches --- proof search versus validation.

        Where tokens are now members of \texttt{Multiset}, the proof of the full formula follows:
        \input{proofnet-example-multiset}
        The above is still tedious and unnatural.
        Proofs of each of $a \vee \neg a$ and $b \vee \neg b$ are constructed in two steps, while the remaining steps are to prove the conjunction of these (proven) subformulae.
        Instead, consider the tokens as members of \texttt{Set} as follows:

        \input{proofnet-example-set}
        This is finally deemed a `satisfying' proof --- dimensionality is extracted as the largest set required in the proof, a nicer algebra of dimensionality may be constructed and there is some essence of justification for the correctness of this $\top$-substitution --- all through changing the datatype of tokens.
    \end{examples}
    
    \begin{remark}
        In essence, allowing for efficient substitution of a $\top$ term for proven subformulae allows the coalescence algorithm to run without restricting the output to additively stratified proofs.
        A proof may be reconstructed including substitutions, where each subproof substituted is itself additively stratified, but now there are contractions for the conclusion of a subproof on top of weakenings for the axioms of a superproof.
    \end{remark}


    \begin{definition*}[Classes of Formulae]
        With this new concept of coalescence, dimensionality and proof search, the definitions in~\ref{sec-ctr:formulae-classes} are revisited and revised.
        Let $\mathcal{D}^i$ be defined as the subclasses of formulae:
        \begin{align*}
            \mathcal{D}^1 \quad &\defeq \quad \top \,|\, \mathcal{D}^1 \wedge \mathcal{D}^1 \,|\, \mathcal{D}^* \vee \mathcal{D}^1 \\
            \mathcal{D}^2 \quad &\defeq \quad P \vee \neg P \,|\, \mathcal{D}^2 \vee \mathcal{D}^{i \geq 2} \,|\, \mathcal{D}^2 \wedge \mathcal{D}^{i \leq 2} \\
            \mathcal{D}^3 \quad &\defeq \quad (P \wedge Q) \vee (P \wedge \neg Q) \vee (\neg P \wedge Q) \vee (\neg P \wedge \neg Q) \,|\, \mathcal{D}^3 \vee \mathcal{D}^{i \geq 3} \,|\, \mathcal{D}^3 \wedge \mathcal{D}^{i \leq 3}
        \end{align*}
        where $\mathcal{D}^* \defeq \bigcup_{i \in \mathbb{N}} \mathcal{D}^i$, $P \in \mathcal{D}^*$, such that $\mathcal{D}^n$ is the class of all formulae provable in exactly n dimensions.
    \end{definition*}
    
    \begin{corollary*}
        For $P \in \mathcal{D}^i, Q \in \mathcal{D}^j$, the following statements hold:
        \begin{equation*}
            P \vee Q \in \mathcal{D}^{\min{i, j}}
            \qquad
            P \wedge Q \in \mathcal{D}^{\max{i, j}}
        \end{equation*}
        In fact, for $P \in \mathcal{D}^n$ in $v$ variables, $n \leq v + 1$.
    \end{corollary*}


    \begin{proposition}[Dimensionality Bounds of Disjunctive Normal Forms]
        Let $\textit{DNF}^n$ denote the set of all formulae in $n$ variables in disjunctive normal form.
        That is, $P \in \textit{DNF}^n \implies P = \bigvee_{j = 1 .. 2^n} (\bigwedge_{i = 1 .. n} a_i)$ the disjunction of $2^n$ conjuctions, each representing an `assignment' of boolean values to $n$ variables.
        Then the following statement holds:
        \begin{equation*}
            P \in \textit{DNF}^n \implies P \in \mathcal{D}^{n+1}
        \end{equation*}
    \end{proposition}

    \begin{proof}
        By induction on the number of variables $n$.

        Let the base case be $n \seteq 1$ and the proof is trivial --- the sequent proof $\vdash a \vee \neg a$ requires 2 dimensions and the hypothesis holds.

        Suppose the inductive hypothesis holds up to $n - 1$ variables --- in $i$ dimensions where $i \leq n$ and all \textit{DNF}-terms in $v$ variables where $i - 1 \equiv v < n$ such that:
        \begin{equation*}
            \forall \,\, i \leq n \,:\, P \in \textit{DNF}^{i} \implies P \in \mathcal{D}^{i}
        \end{equation*}
        We seek to show the hypothesis continues to hold for up to $n$ variables:
        \begin{equation*}
            \forall \, Q \in \textit{DNF}^n :  Q \in \mathcal{D}^{n+1}
        \end{equation*}
        Let $a_n$ be fresh in $P$ and $a_1 \ldots a_{n-1}$ be used, then the disjunctive normal form expression $P$ is in the class:
        \begin{equation*}
            P \defeq * \,|\, P_j \vee P    \qquad    P_j \defeq * \,|\, a_i \,|\, \neg a_i \,|\, P_j \wedge P_j
        \end{equation*}
        Write $P \defeq \bigvee_j (P_j \defeq \bigwedge_j a_{ij})$ for some $P_j \,\, \forall \, j \in 1 \ldots n$ where $P_j$ is purely a conjunction of an `assignment' $a_{ij}$ of variables.
        Now construct $Q$ as follows:
        \begin{equation*}
            Q \equiv Q_{l} \vee Q_{r} \seteq (\bigvee_i P_i \wedge a_n) \vee (\bigvee_i P_i \wedge \neg a_n) \in \textit{DNF}^n
        \end{equation*}
        where $Q$ is the simplest $\textit{DNF}^n$-term extension of $P$.
        Each $P_i$ is of course not provable by itself, therefore not provable up to $n$ dimensions, but up to $n$ dimensions produces sequents of $n$ terms.
        Similarly, $Q$ is therefore not provable in $n$ dimensions.
        
        Since $Q$ not provable in $n$ dimensions, extend the proof to $n+1$ dimensions --- on the left, the leaf sequents of $Q_l$ may be extended from $\vdash \Gamma$ atomic terms to $\vdash \Gamma, \neg a$ giving $n + 1$ dimensionality as required.
        Now the conclusion of this proof subtree may be up to:
        \begin{prooftree}
            \AxiomC{}
            \doubleLine\UnaryInfC{$\vdash Q_l \ldots_n Q_l, \neg a_n$}
            \UnaryInfC{$\vdash Q \ldots_n Q, \neg a_n$}
        \end{prooftree}
        Similarly, this may be applied to $Q_r$ but instead `solving' for each $a_{i < n}$ yielding $\,\, \forall \, i < n$:
        \begin{prooftree}
            \AxiomC{}
            \doubleLine\UnaryInfC{$\vdash Q_r \ldots_n Q_r, a_i$}
            \UnaryInfC{$\vdash Q \ldots_n Q, a_i$}
        \end{prooftree}
        
        Reducing across all subtrees by the $\wedge$ rule yields $\vdash Q \ldots_n Q, P_j \wedge a_n$ for some assignment $P_j$.
        Finally, the $\vee$ rule may be applied up the formula syntax tree to yield $\vdash Q \ldots_{n} Q, Q \equiv \, \vdash Q \ldots_{n+1} Q \implies Q \in \mathcal{D}^{n+1}$, where $Q \in \textit{DNF}^n$ as required.
    \end{proof}

    \begin{remark} \label{remark:proof-subtleties}
        This proof omits the subtleties and specifics of deriving each assignment of $P_j$ with respect to the two cases of $a_n, \neg a_n$.
        This is considered tedious and does not aid to a clear representation of the intermediate steps but remains to be formally proven, should such a thing be desirable.
    \end{remark}

    \begin{proposition*}
        
        \begin{equation*}
            P \in \mathcal{D}^{n+1} \iff \exists \, Q \in P \text{ such that } Q \in DNF^{n}
        \end{equation*}
    \end{proposition*}

    
    \begin{remarks}
        \citet{tableaux-for-logic-of-proofs}
        \citet{connection-based-proof-method}
    \end{remarks}
